---
title: "p8105_hw2_SZ3319"
author: "Shiyu Zhang"
date: "2024-09-30"
output: github_document


---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(haven)
library(tidyr)
```

## Problem 1

### 1.1 Import the dataset and clean names

```{r, message = FALSE}
transit_df = 
  read_csv("hw2data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na = c("NA", "", ".")) |> 
  janitor::clean_names()

tail(transit_df, 10)
```

### 1.2 Select variables

Select columns and clean the data; convert `entry` to logical.

```{r, message = FALSE}
transit_select <- transit_df |> 
  select(
    line, 
    station_name, 
    station_latitude, 
    station_longitude, 
    starts_with("route"), 
    entry, 
    vending, 
    entrance_type, 
    ada) |> 
  mutate(
    entry = ifelse(entry == "YES", TRUE, FALSE)
    )
  
```

Description:



### 1.3 Analysis the data

#### 1.3.1 Calculate the distinct station.

```{r, message = FALSE}
distinct_stations <- transit_select |> 
  distinct(line, station_name) |> 
  nrow()

distinct_stations
```
There are **465** distinct stations are there.

#### 1.3.2 Calculate the number of ADA compliant.

```{r}
ADA_compliant_stations <- 
  transit_select |> 
  filter(ada == "TRUE") |> 
  distinct(line, station_name) |> 
  nrow()

ADA_compliant_stations
```

There are **84** ADA compliant stations.

#### 1.3.3 Propotion of station

Calculate the total number of stations without vending.

```{r, message = FALSE}
no_vending <- transit_select |> 
  filter(vending == "NO") |>
  nrow()
```
 
Calculate how many of them allow entry.

```{r, message = FALSE}
allow_entry_no_vending <- transit_select |> 
  filter(vending == "NO", entry == TRUE) |>
  nrow()
```

Calculate the proportion.

```{r}
proportion_no_vending <-
  allow_entry_no_vending / no_vending

proportion_no_vending
```
The proportion of station entrances without vending allow entrance is **0.3770492**.

### 1.4 Reform the data.

```{r}

```


## Problem 2


### 2.1 Import the `Mr. Trash Wheel` sheet and clean names

```{r, message = FALSE}
mr_df = 
  read_excel("hw2data/202409 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", skip = 1) |> 
  janitor::clean_names()

head(mr_df, 10)


mr_clean =
  mr_df |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster) |> #omit rows that do not include dumpster-specific data
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    year = as.numeric(year),
    trash_wheel = "Mr"
  ) # round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

tail(mr_clean, 10)
```


### 2.2 Import the `Professor Trash Wheel` sheet and clean names

```{r}
professor_df = 
  read_excel("hw2data/202409 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", skip = 1) |> 
  janitor::clean_names()

head(professor_df, 10)


professor_clean =
  professor_df |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster, month) |> #omit rows that do not include dumpster-specific data
  mutate(trash_wheel = "Professor")
  

tail(professor_clean, 10)
```


### 2.3 Import the `Gwynnda Trash Wheel` sheet and clean names

```{r}
gwynnda_df = 
  read_excel("hw2data/202409 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", skip = 1) |> 
  janitor::clean_names()

head(gwynnda_df, 10)


gwynnda_clean =
  gwynnda_df |> 
  select(dumpster:homes_powered) |> 
  drop_na(dumpster) |> 
  mutate(trash_wheel = "Gwynnda")

tail(gwynnda_clean, 10)
```


### 2.4 Bind the datasets

```{r}
bind_data = 
  bind_rows(mr_clean, professor_clean, gwynnda_clean) |>
  janitor::clean_names() |>
  select(trash_wheel, everything()) 

head(bind_data, 10)
```
### 2.5 Total description

```{r}
total_weight_professor =
  bind_data |> 
  filter(trash_wheel == "Professor") |> 
  summarise(total_weight = sum(weight_tons, na.rm = TRUE)) |> 
  pull(total_weight)

total_weight_professor
```

```{r}
total_cigarette_butts_gwynnda =
  bind_data |> 
  filter(trash_wheel == "Gwynnda" & year == 2022 & month == "June") |> 
  summarise(
    total_cigarette_butts = sum(cigarette_butts, na.rm = TRUE)
  ) |> 
  pull(total_cigarette_butts)

total_cigarette_butts_gwynnda
```



## Problem 3

### 3.1 Import data and clean names

Import the data and clean names.

```{r}
baker_df = 
  read_csv("hw2data/gbb_datasets/bakers.csv") |> 
  janitor::clean_names()

bake_df =
  read_csv("hw2data/gbb_datasets/bakes.csv") |> 
  janitor::clean_names()

result_df = 
  read_csv("hw2data/gbb_datasets/results.csv", skip = 2) |> 
  janitor::clean_names()

viewer_df =
  read_csv("hw2data/gbb_datasets/viewers.csv") |> 
  janitor::clean_names()

```

**bakers**: Separate the bakers' name in `bakers` dataset.

```{r}
baker_clean = 
  baker_df |> 
  separate("baker_name", into = c("baker", "last_name"), sep = " ") |> 
  distinct() 
```

**bakes**: Exchange `Jo` to `"Jo"`.

```{r, message = FALSE}
bake_clean = 
  bake_df |> 
  distinct() |> 
  mutate(baker = ifelse(baker ==  '"Jo"', "Jo", baker))
```

### 3.2 Set up a dataset combined with data from `bakers.csv`, `bakes.csv`, `results.csv`.

Check if all bakers are there in `bake_clean`, `result_df` and `baker_clean`.

```{r, message = FALSE}
ms1_data = anti_join(result_df, baker_clean, by = "baker")

ms1_data
```
Joanne is shown in `result`, but not in `bakers`.

```{r, message = FALSE}
ms2_data = anti_join(
  bake_clean, baker_clean,
  by = c("baker"))

ms2_data
```
All bakers in `bake` are shown in `baker`.

Combine the datasets.
```{r, message = FALSE}
combined_df = 
  baker_clean |> 
  left_join(bake_clean, 
            by = c("baker", "series")) |> 
  left_join(result_df, 
            by = c("baker", "series", "episode")) |> 
  select(-(last_name)) |> 
  arrange(series, episode, technical) |> 
  select(series, episode, baker, everything())
  
```



### 3.3 Export the `combined_df`.

```{r export_data}
write_csv(combined_df,
          "hw2data/gbb_datasets/combined_df.csv")
```

The final `combined_df` dataset has **573** observations and **10** variables. 

### 3.4 Table for readers

```{r}

```


